Practical Machine Learning course project
=====================

In this project I've analyzed data to determine quality of activivty of people who lifted a barbell. People for this study makes mistakes in exercises and goal of this project is to prdeict manner they made exercises.

For reproducing this script you should install caret package and randomForst package.

Getting the data
-----------------

Download data if its needed:

```{r}
if (!file.exists("pml-testing.csv")){
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", "pml-testing.csv", method="curl")
}

if (!file.exists("pml-training.csv")){
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", "pml-training.csv", method="curl")

}
```

Reading data:

```{r}
testing_data <- read.csv("pml-testing.csv", na.strings=c("#DIV/0!"))
training_data <- read.csv("pml-training.csv", na.strings=c("#DIV/0!"))
```

We're interested in numeric variables so make it numeric:
```{r}
for(i in c(8:ncol(training_data)-1)) {training_data[,i] = suppressWarnings(as.numeric(as.character(training_data[,i])))}

for(i in c(8:ncol(testing_data)-1)) {testing_data[,i] = suppressWarnings(as.numeric(as.character(testing_data[,i])))}
```

According to summary of data there are few columns with only na meanings. Also we're not interested in first 7 columns contains data about time and person who was under experiment. They will not help for analyzis so lets drop it:

```{r}
clean_training_data <- training_data[,colSums(is.na(training_data)) == 0][-(1:7)]

clean_testing_data <- testing_data[,colSums(is.na(testing_data)) == 0][-(1:7)]
```

Training and cross-validation sets
-----------------------------------
So we has create clean datasets and ready to analyze data.

Let's separate training set and cross validation set to be sure our model works fine.
```{r}
library(caret)
set.seed(1000)
cl <- createDataPartition(y=clean_training_data$classe, p=0.7, list=FALSE)
training_set <- clean_training_data[cl,]
cv_set <- clean_training_data[-cl,]

```

Random Forest
-------------
For analysis was chosen Random Forest algorithm.

```{r}
library(randomForest)
r_f <- randomForest(classe ~. , data<- training_set)
```
predictions_training <- predict(r_f, newdata=training_set)
predictions_cv <- predict(r_f, newdata=cv_set)
confusionMatrix(predictions_cv, cv_set$classe)

As we can see from confusion matrix, random forest gives good results on validation sets.

Submission assigment
--------------------

The last part of this project is evaluation on test set we were given at the beginning.
According to instructions we should use our prediction algorythm on testing set and create 20 files with class column.

```{r}
predictions_testing <- predict(r_f, newdata=clean_testing_data)

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(predictions_testing)
```

System was trained perfectly and all 20 cases were decided right.
